\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-11-03 demystify.tex
updates

\section{Summary}\label{sec:summary}
Here we summarize the main points of the document. This allows you to quickly skim through the document and decide which parts you want to read in detail.
\begin{enumerate}
\item Should you apply to a PhD in CS in the US?
\begin{itemize}
    \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\S\ref{sec:should}).
\end{itemize}
\item How is your application evaluated?
\begin{itemize}
    \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically 3 faculty (\S\ref{sec:evalapps}).
    \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance. However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself (\S\ref{sec:contact}).
\end{itemize}
\item Application Materials
\begin{itemize}
    \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, personal statement, and recommendation letters.
    \item LORs are important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
    \item GRE \emph{is not} required by most schools (\S\ref{sec:grades}).
    \item Grades are important, but only if your school is well-known (\S\ref{sec:grades}).
    \item Personal statement is very important. Write it in such a way that makes you \emph{stand out} (\S\ref{sec:research-statement} and \S\ref{sec:improve-your-chance})
    \item Getting an interview is typically a \emph{good sign} (no one interviews weak candidates).
\end{itemize}
\item What to do after getting admitted?
\begin{itemize}
    \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\S\ref{sec:accepted}).
\end{itemize}
\item Funding
\begin{itemize}
    \item TA and RA are two main sources of funding.  TA (teaching assistantship) is provided by the department to help profs. with classes (e.g., grading). RA (research assistantship) is provided by profs. to help them with their research (\S\ref{sec:funding}).
\end{itemize}
\item Choosing School and Professors
\begin{itemize}
    \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} be advisor, i.e., cannot graduate PhD students  (\S\ref{sec:schoolsandprofs}).
    \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\S\ref{sec:contact}).
\end{itemize}
\item Miscs and FAQS
\begin{itemize}
    \item Increasing your admission chance by being unique and standing out (\S\ref{sec:improve-your-chance}).
    \item You can successfully apply to CS PhD even if you have non-STEM background (\S\ref{sec:non-stem}).
    \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\S\ref{sec:non-us-differences}).
    \item Your PhD costs about \$400K in total, but you do not pay for it (\S\ref{sec:ra-cost}).
    \item Despite some miserable stories on social media, many PhD students have good mentors, supportive labmates, healthy working environment ... and are happy (\S\ref{sec:happy}).
\end{itemize}
\end{enumerate}

@@

\paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc. Even if you can't come in person, you still can attend virtually and meet with individual faculty.
During the Open House, you get a chance to talk to individual faculty and current students.  Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).  Talk to students about their advisors, the dept, the area, funding situation etc.  Ask about anything you want to determine that they deserve \emph{you}.

@@

\subsubsection{Teaching Assistant (TA)}

TA is common in the beginning when you haven't found your advisor who would pay you RA. As a TA, you spend up to 20 hrs/week and help professors with their classes (e.g., grading or teaching labs/recitation). Your TA is paid through the school/department, i.e., they hire you to help teach.  During a semester, a TA might work with several courses and professors (not necessary their advisor).  TA funding typically is not available during the summer, which has much fewer courses.

@@

\textbf{How to get RA?} When a professor recruits you, they will likely give you RA right away (e.g., when you apply).  A common scenario is that you first get admitted with TA, and then after a year or two find an advisor to support you with RA. See how to contact a prof. in \S\ref{sec:contact} for research opportunities.

@@

Choosing a school and an advisor is clearly one of the most important things in your mind when you apply to and especially when you get admitted.  This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.  This section aims to mitigate some confusions and help you make informed decision.

@@

Now we discuss about finding advisors.

\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
There are several types of faculty in a university that has a CS PhD program.  You might hear about tenured,  tenure track, teaching faculty.  You might also things such as assistant, associate, full, adjunct, emeritus, teaching, research professors. Here is a primer on these terminologies.

\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, mentoring Ph.D students.  They often have very low teaching load (e.g., 1 per semester). \emph{Tenured} faculty are professors who have been promoted to have a permanent position (informally, very hard to fire them).   \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.  Both tenure and tenured track faculty can serve as a formal advisor of Ph.D. students . This is very important because you need to have a tenure line faculty as your main PhD advisor. \S\ref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your advisor.

In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching. They typically teach 3-4 classes per semester and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers, obtaining grants).  They can mentor Ph.D. students but they typically cannot serve as a formal advisor of Ph.D. students (i.e., they \emph{cannot graduate} PhD students).

\paragraph{Rankings} Faculty have ranking, e.g., assistant, associate, full, regardless if they are tenure line or teaching.  Assistant means new faculty, associate means they have been promoted, and full means they are senior. A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).

Adjunct means they are not full-time faculty, e.g., they might be working in industry and teach a class or two.  Emeritus means they are retired but still have some affiliation with the university.  Research faculty (or scientist) are typically non-tenure line faculty who focus on research.  Due to their roles, adjunct, emeritus, and research faculty typically do not serve as a formal advisor of Ph.D. students.

@@

You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.
Ultimately, choose one that fits you the most by communicating with them, meeting them, asking them questions, and even talking to their current students.

\begin{commentbox}[Thanh:]
    In my opinion, having a well-suited advisor is crucial for a successful PhD and research career. One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies. An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs. Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities. Unfortunately, VinAI's resident program is limited to AI research.

    In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants? \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
    \end{commentbox}

\begin{commentbox}[Xiaokuan:]
Whether the student's research interest matches that of the advisor is very important;
if there is a mismatch,
either the student or the advisor has to make compromises,
which often leads to disagreements or conflicts. 
IMO, the advisor should be the one who {\it guides}  students to do research while allowing students to pursue their own interest,
instead of {\it dictating} their research.

\end{commentbox}
% \subsubsection{Strong Faculty vs High-ranked School}
% \begin{commentbox}[Thanh:]
% When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation? Of course, both are great, but it is hard to achieve both.  I believe receiving some guidance from you  would be incredibly valuable.
% \end{commentbox}

@@

\begin{commentbox}[Xiaokuan]
    Applying for Ph.D. and contacting a potential Ph.D. advisor is a classic {\bf `why me, why you'} problem, 
similar to looking for a job in a company.
On a high level,
you need to show that you have done your {\it homework}
regarding the professor and the university,
and clearly explain:
1) why do you think you are a good fit in professor A's group?
2) why do you want to be advised by professor A, not B? 
3) why do you want to apply for university X, not Y?
If you don't want to spend time to do your homework,
the chance of getting a reply is close to zero.
\end{commentbox}

@@

However, if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty. In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} advisor of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main advisor will need to be a tenured or tenure-track faculty in CS. If in doubt, you should check with the CS department for their requirements.  For this specific reason,  CSRankings includes only tenured or tenured track faculty who can advise CS PhD students. I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.

@@

\emph{Project proposal}: in many countries, you have to choose a project and advisor \emph{during} the application process (e.g., you write a proposal to a potential advisor). In the US, you often start your PhD without an advisor or project and find them later. This is also why US PhD programs are longer.

\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential advisor and research topics. In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you proposed with the advisor you chose. Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}. In other countries, you do not have to do these exams or only do few of them.

\emph{Funding}:  In Europe, stipend (salary) comes from the university or from gov't research organizations and often have a fixed duration (e.g., 3 years).  In the US, stipend (e.g., RA) comes directly from your advisor (no fixed duration).  There are also fewer TA opportunities in the European universities compared to the US.

@@

\subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
PhD students often ask why their salary is so low compared to ludicrous grants their advisors get. They also wonder why their offer letters sometime mentioned that their benefits appear better than what they actually receive (i.e., stipend).  This section aims to shed some light to these questions.

@@

At the end, the total budget comes out to be \$69K/year to support a PhD student. \emph{The summary is that over your 5-6 year of your PhD, you cost about \$350K--400K, and while you're paid X, your advisor probably pays 2X for you}.

@@

\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US). However, \emph{do not} install or use them on university computers (e.g., those given by your department or advisor).  It is unlikely that the university will track you down, but it is the \emph{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or advisor being sued for a large sum of money, and it is \emph{you} who caused it.  If you need software, ask your advisor or the department.

>>>>

\section{Summary}\label{sec:summary}
Here we summarize the main points of the document. This allows you to quickly skim through the document and decide which parts you want to read in detail.
\begin{enumerate}
\item Should you apply to a PhD in CS in the US?
\begin{itemize}
    \item \emph{Yes, definitely}.  CS PhD study in the US is fully funded and admission into good universities is not any harder than non-US schools (\S\ref{sec:should}).
\end{itemize}
\item How is your application evaluated?
\begin{itemize}
    \item Applications are evaluated by the \emph{PhD Admission} committee and each application is reviewed by typically 3 faculty (\S\ref{sec:evalapps}).
    \item Individually faculty \emph{cannot directly admit} a student---so do not email and ask if you have a chance. However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself (\S\ref{sec:contact}).
\end{itemize}
\item Application Materials
\begin{itemize}
    \item The committee will look at various factors, but the most important ones are research ability, e.g., publications, personal statement, and recommendation letters.
    \item LORs are important, but only if they are personalized and talk about your research ability (\S\ref{sec:lor}).
    \item GRE \emph{is not} required by most schools (\S\ref{sec:grades}).
    \item Grades are important, but only if your school is well-known (\S\ref{sec:grades}).
    \item Personal statement is very important. Write it in such a way that makes you \emph{stand out} (\S\ref{sec:research-statement} and \S\ref{sec:improve-your-chance})
    \item Getting an interview is typically a \emph{good sign} (no one interviews weak candidates).
\end{itemize}
\item What to do after getting admitted?
\begin{itemize}
    \item Attend \emph{Open House} to learn more about the place and \emph{interview} profs---they would be much more willing to talk to you now (\S\ref{sec:accepted}).
\end{itemize}
\item Funding
\begin{itemize}
    \item TA and RA are two main sources of funding.  TA (teaching assistantship) is provided by the department to help profs. with classes (e.g., grading). RA (research assistantship) is provided by profs. to help them with their research (\S\ref{sec:funding}).
\end{itemize}
\item Choosing School and Professors
\begin{itemize}
    \item Many schools do not offer PhD studies in CS and many CS professors \emph{cannot} be adviser, i.e., cannot graduate PhD students  (\S\ref{sec:schoolsandprofs}).
    \item Contacting a prof. for research opportunities is recommended, but do it \emph{properly} (\S\ref{sec:contact}).
\end{itemize}
\item Miscs and FAQS
\begin{itemize}
    \item Increasing your admission chance by being unique and standing out (\S\ref{sec:improve-your-chance}).
    \item You can successfully apply to CS PhD even if you have non-STEM background (\S\ref{sec:non-stem}).
    \item Compared to other countries, CS PhD in the US does not require an MS degree but has longer PhD study time (\S\ref{sec:non-us-differences}).
    \item Your PhD costs about \$400K in total, but you do not pay for it (\S\ref{sec:ra-cost}).
    \item Despite some miserable stories on social media, many PhD students have good mentors, supportive labmates, healthy working environment ... and are happy (\S\ref{sec:happy}).
\end{itemize}
\end{enumerate}

@@

\paragraph{Open House} Most schools will have \textbf{Open House} events, which are a great resource to learn about the school, department, faculty, research, living, etc. Even if you can't come in person, you still can attend virtually and meet with individual faculty.
During the Open House, you get a chance to talk to individual faculty and current students.  Take notes of faculty who make you excited, count those that are taking in new students (if they meet you, likely they are considering new students!).  Talk to students about their advisers, the dept, the area, funding situation etc.  Ask about anything you want to determine that they deserve \emph{you}.

@@

\subsubsection{Teaching Assistant (TA)}\label{sec:ta}

TA is common in the beginning when you haven't found your adviser who would pay you RA. As a TA, you spend up to 20 hrs/week and help professors with their classes (e.g., grading or teaching labs/recitation). Your TA is paid through the school/department, i.e., they hire you to help teach.  During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding typically is not available during the summer, which has much fewer courses.

@@

\textbf{How to get RA?} When a professor recruits you, they will likely give you RA right away (e.g., when you apply).  A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA. See how to contact a prof. in \S\ref{sec:contact} for research opportunities.

@@

Choosing a school and an adviser is clearly one of the most important things in your mind when you apply to and especially when you get admitted.  This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.  This section aims to mitigate some confusions and help you make informed decision.

@@

Now we discuss about finding advisers.

\subsubsection{Finding the right adviser}
Obviously, there is no one-size-fit-all answer to this question. The best adviser is the one that you can work well with, i.e., \emph{fits} you.  This is difficult and takes time.  Fortunately, unlike many countries that require finding an adviser and research topic before starting the PhD (\S\ref{sec:non-us-differences}), CS PhD programs in the US will typically give you a couple of years to search for advisers and research topics.  This is especially true if you're admitted with TA (\S\ref{sec:ta}), which gives you time to explore and find the right adviser.

\begin{commentbox}[Xiaokuan:]
    Whether the student's research interest matches that of the adviser is very important;
    if there is a mismatch,
    either the student or the adviser has to make compromises,
    which often leads to disagreements or conflicts. 
    IMO, the adviser should be the one who {\it guides}  students to do research while allowing students to pursue their own interest,
    instead of {\it dictating} their research.
    \end{commentbox}

Here are some general advice to find professor.  First, search for profs. that share similar research interests. You can learn about faculty and their level of research activity through the faculty's website or even from CSRankings.  For example, in CSRankings, if you want to work with PL, you can search for those publish in PL conferences.  If you want to work with SE \emph{and} AI, you can search for faculty who work in both SE and AI.   You can also go to their website, look at their papers and projects.  In any case, you should contact the professor and talk to them. For example, you can (cold) email them and ask about their research and if they are taking new students and schedule appointments to chat with them (office hrs are often the best time).  See \S\ref{sec:contact} for more details on contacting professors.

Another very effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore around).  Professors teaching these \emph{special topics courses} and \emph{research seminars} typically are recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the profs after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).  
You can also ask if you can do an independent study or research with a prof. This can be informal (no credit) and takes place during the summer or winter break.  For example, I work with multiple students, many of whom are undergrads, this way. Many will drop because they find they don't like my research, but some will stay and become my PhD students.

Ultimately, choose one that fits you the most by communicating with them, taking their courses, meeting them, asking them questions, and even talking to their current students. Indeed all this takes a lot of time and effort, but remember, you will be working with this person for 4-5 years, so it is important to try to find the right one.

\begin{commentbox}[Thanh:]
    In my opinion, having a well-suited adviser is crucial for a successful PhD and research career. One effective approach to finding a suitable professor is by working with a professor during your undergraduate studies. An exemplary instance is VinAI's residency program, where residents collaborate with professors from the US for two years before applying to PhD programs. Many VinAI residents have achieved remarkable results and gained admission to prestigious US universities. Unfortunately, VinAI's resident program is limited to AI research.

    In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants? 
    \tcblower
    \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities. Many probably will say no as they do not have the bandwidth to take on random strangers, but some may say yes if they see you have a matching profile.  
    \end{commentbox}

\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
Not every faculty in the CS program can serve as your formal adviser. Let's try to understand the different types of faculty and their roles. For example, you might hear about tenured,  tenure track, teaching faculty.  You might also things such as assistant, associate, full, adjunct, emeritus, teaching, research professors. Here is a primer on these terminologies.

\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, mentoring Ph.D students.  They often have very low teaching load (e.g., 1 per semester). \emph{Tenured} faculty are professors who have been promoted to have a permanent position (informally, very hard to fire them).   \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.  \textbf{Both tenure and tenured track faculty can serve as a formal adviser of Ph.D. students}. This is very important because you need to have a tenure line faculty as your main PhD adviser. \S\ref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.

In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching. They typically teach 3-4 classes per semester and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers, obtaining grants).  They can mentor Ph.D. students but they typically cannot serve as a formal adviser of Ph.D. students (i.e., they \emph{cannot graduate} PhD students).

\paragraph{Titles} Faculty have titles, e.g., assistant, associate, full, regardless if they are tenure line or teaching.  Assistant means new faculty, associate means they have been promoted, and full means they are senior. A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).

Adjunct faculty not full-time faculty, e.g., they might be working in industry and teach a class or two for fun.  Emeritus means they are retired but still have some affiliation with the university.  Research faculty (or scientist) are typically non-tenure line faculty who focus on research.  Due to their roles, adjunct, emeritus, and research faculty typically do not advise Ph.D. students.

@@

% \subsubsection{Strong Faculty vs High-ranked School}
% \begin{commentbox}[Thanh:]
% When considering PhD programs, we often wonder if we should prioritize a high-ranking university or a professor with a strong reputation? Of course, both are great, but it is hard to achieve both.  I believe receiving some guidance from you  would be incredibly valuable.
% \end{commentbox}

@@

\begin{commentbox}[Xiaokuan]
    Applying for Ph.D. and contacting a potential Ph.D. adviser is a classic {\bf `why me, why you'} problem, 
similar to looking for a job in a company.
On a high level,
you need to show that you have done your {\it homework}
regarding the professor and the university,
and clearly explain:
1) why do you think you are a good fit in professor A's group?
2) why do you want to be advised by professor A, not B? 
3) why do you want to apply for university X, not Y?
If you don't want to spend time to do your homework,
the chance of getting a reply is close to zero.
\end{commentbox}

@@

However, if your intention is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty. In fact, if a faculty is not in CS, it is unlikely that they can be the \emph{main} adviser of a CS PhD student---they may \emph{co-advise} or be in the PhD dissertation committee, but your main adviser will need to be a tenured or tenure-track faculty in CS. If in doubt, you should check with the CS department for their requirements.  For this specific reason,  CSRankings includes only tenured or tenured track faculty who can advise CS PhD students. I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students.

@@

\emph{Project proposal}: in many countries, you have to choose a project and adviser \emph{during} the application process (e.g., you write a proposal to a potential adviser). In the US, you often start your PhD without an adviser or project and find them later. This is also why US PhD programs are longer.

\emph{Course work}: in the US you will spend the first couple of years taking classes and explore potential adviser and research topics. In other countries, you often do not do course work and start your research right away, e.g., you immediately work on the research project you proposed with the adviser you chose. Also, in the US you also have to pass a series of "exams", e.g., qualifying exam, comprehensive exam, thesis proposal defense\footnote{The word "ABD" (all but dissertation) is used in the US to refer to a PhD candidate who have finished all course work and exams and only need to write and defend their dissertation.}. In other countries, you do not have to do these exams or only do few of them.

\emph{Funding}:  In Europe, stipend (salary) comes from the university or from gov't research organizations and often have a fixed duration (e.g., 3 years).  In the US, stipend (e.g., RA) comes directly from your adviser (no fixed duration).  There are also fewer TA opportunities in the European universities compared to the US.

@@

\subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
PhD students often ask why their salary is so low compared to ludicrous grants their advisers get. They also wonder why their offer letters sometime mentioned that their benefits appear better than what they actually receive (i.e., stipend).  This section aims to shed some light to these questions.

@@

At the end, the total budget comes out to be \$69K/year to support a PhD student. \emph{The summary is that over your 5-6 year of your PhD, you cost about \$350K--400K, and while you're paid X, your adviser probably pays 2X for you}.

@@

\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries (and even in the US). However, \emph{do not} install or use them on university computers (e.g., those given by your department or adviser).  It is unlikely that the university will track you down, but it is the \emph{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or adviser being sued for a large sum of money, and it is \emph{you} who caused it.  If you need software, ask your adviser or the department.
